---
title: "create_project"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{create_project}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(dplyr)
```

## Introduction

This vignette will walk you through the `create_project()` function.
The function creates a new directory or populates an existing project directory with essential files needed for a new analysis project.
Existing files in the directory will not be overwritten, unless the user specifically requests to overwrite them. 

By default the directory is populated with customized `README.md`, `.gitignore`, and `<name>.Rproj` files.
Users can also create a personalized project templates. Instructions for creating your own template is detailed in the [custom templates section](#custom_template). 

## Default Template

Let's create a new project and inspect the results.
The new project will be created in a the folder called `"My Project Folder"`.

```{r, message=FALSE}
# loading packages
library(starter)

# specifying project folder location (folder does not yet exist)
project_path <- fs::path(tempdir(), "My Project Folder")
```

Let's set up our new project. 

```{r, comment=""}
create_project(
  path = project_path,
  open = FALSE # don't open project in new RStudio session
)
```

The directory was created, files added, a git repository was initialised in the folder, and the reproducible environment was constructed with the renv package.

### **`README.md`**

The following is the `README.md` file added the project directory.
It's pre-filled with information specific to this project.
The first line of the file is the project folder name, followed by a R code for creating a symbolic link from the project folder to the secure data folder.

```{r, comment = "", echo=FALSE}
readr::read_file(fs::path(project_path, "README.md")) %>% cat()
```

### **`.gitignore`**

The `.gitignore` file is populated with various R files, data file types, and miscellaneous files typically not wanted in a Git repository.

```{r, comment = "", echo=FALSE}
readr::read_file(fs::path(project_path, ".gitignore")) %>% cat()
```

## Custom Template {#custom_template}

While the default template is useful, it can, of course, be useful to create a custom template for yourself or your team.
To create a custom template, you'll need two things: 

1. Template files

1. Meta data for each template file (e.g. location of template file, file name, etc.)

### Template Files

To begin, create a folder to save all the template files in.
There are two kinds of template files.
The first are static files that a merely copied into the new project folder.
These files contain no custom information about your new project.
Save these files in the template folder.
The `.gitignore` file described above in the default template is an example of a file static file.

The second type of template contains data or information specific to the new project.
In the default template described above, the `README.md` file is an example of these dynamic template files.
The `README.md` file is populated with the project name, and the project-specific code for creating a symbolic link to the secure data folder.
The following fields are available to include in the dynamic templates.

    {{folder_name}}    project folder name
    {{symbolic_link}}  code for establishing symbolic link to data folder
    {{foo()}}          any R function, e.g. `Sys.Date()`

Each of these fields are accessible via standard `glue::glue()` syntax, with the exception that double curly brackets are needed rather than single (e.g. `glue::glue(., .open = "{{", .close = "}}")`).

Below is an example of the dynamic `README.md` template from the default template.

```{r, comment = "", echo=FALSE}
readr::read_file(system.file("project_templates/default_readme.md", package = 'starter')) %>% cat()
```

### Metadata

In order to implement and use a template, you must create a metadata object that stores information about each template file.
The metadata is stored as a named list: one element in the list per template file.

It is recommended you create the list elements named 'readme' and 'gitignore'. 
The 'readme' element is used in the in the `use_project_readme()` function and 'gitignore' is in the `use_project_gitignore()` function---without these elements, these useful functions will return errors.

The list elements may be named any proper name within R.
However, the names must be unique and we recommend short names.
These do not need to match/reference any other text in the template file.

```{r}
my_template <-
  quote(list(
    gitignore = list(),
    readme = list(),
    rproj = list()
  ))
```

Each element of the list is itself a list.
The list is quoted so the template meta data is not evaluated until you call the `create_project()` function, allowing files to be dynamically named and template files to be saved within a package.
The list contains three important pieces of information about each template file.

1. The template path and file name.

1. The name of the file after it's been copied to the new project folder.

1. A logical argument indicating whether the file is a straight copy, or a dynamic file (described above).

The metadata for the `README.md` file in the default template looks like this:

```{r, eval=FALSE}
readme <- list(
  template_filename = system.file("project_templates/default_readme.md", package = "starter"), 
  filename = "README.md", 
  glue = TRUE
)
```

As another example, here's the metadata for the `*.Rproj` file from the default template.
Note that the file name can be dynamic (the use of `glue::glue()`), while the underlying file is not dynamic (`glue = FALSE`).

```{r, eval=FALSE}
rproj <- list(
  template_filename = system.file("project_templates/default_rproj.Rproj", package = "starter"), 
  filename = glue::glue("{folder_name}.Rproj"), 
  glue = FALSE
)
```

The final metadata object is a quoted list of each of the individual file metadata objects.
It's important to note that this is a quoted list!
The list cannot be evaluated until the new project is created.

```{r}
my_template <- quote(list(
  readme = list(
    template_filename = system.file("project_templates/default_readme.md", package = "starter"), 
    filename = "README.md",
    glue = TRUE
  ),
  rproj = list(
    template_filename = system.file("project_templates/default_rproj.Rproj", package = "starter"), 
    filename = glue::glue("{folder_name}.Rproj"),
    glue = FALSE
  ),
  
gitignore = list(
  template_filename = system.file("project_templates/default_gitignore.txt", package = "starter"), 
  filename = "gitignore.txt", 
  glue = TRUE
)
))
```

You may also `source()` an R script by adding the path to the file as a template attribute.
The script will be sourced after project template has been placed.

```r
attr(my_template, "script_path") <- "<path to file>"
```

There is one last step---give your template a label.
The label will be displayed each time the template is used in either `create_project()` or `use_project_file()`.

```{r}
attr(my_template, "label") <- "My Very 1st Project Template"
```

### Implement Custom Template

Now that you have a folder with your template files stored and you have created a metadata list object, you can use them to create a new project folder with your custom template.
The following code will get you started.

```{r, include=FALSE, echo=FALSE}
unlink(project_path, recursive = TRUE)
```

```{r}
create_project(
  path = project_path,
  template = my_template # metadata list created above
)
```
